---
title: "Marvel Rivals Exploration"
subtitle: "Drafting Vizualization - EDS 240"
author: "Marie Tolteca"
date-modified: last-modified
format: html
---
1. Restate the questions you hope to answer with your inforgraphic. This should include one overarching question (think of this as driving the overall theme of your infographic) and at least three subquestions (each of which will be addressed by your infographic’s component visualizations). Have these questions changed at all since FPM #1? If yes, how so? 

- Overall Question: 
- 

```{r}
library(tidyverse)
library(here)
library(dplyr)
library(jsonlite)
library(httr)
library(RCurl)
library(tibble)
library(ggimage)
```


```{r}
# Progress by hero across seasons
hero_season_progress <- clean_season %>%
  group_by(hero_name, season) %>%
  summarise(
    games = n(),
    win_rate = mean(win_rate == "Win")
  ) %>%
  filter(games > 2)  # Only heroes with 3+ games per season

# Heatmap version (shows all at once)
ggplot(hero_season_progress, aes(x = factor(season), y = hero_name, fill = win_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "red", high = "green") +
  labs(x = "Season", y = "", fill = "Win Rate",
       title = "Hero Performance Heatmap")
```

# Average K/D By Role
```{r}
role_kd_summary <- clean_season %>%
  group_by(role) %>%
  summarise(
    total_kills = sum(kills),
    total_deaths = sum(deaths),
    avg_kills = mean(kills),
    avg_deaths = mean(deaths),
    kd_ratio = sum(kills) / sum(deaths),
    games = n()
  )

role_kd_summary

# Side-by-side kills vs deaths
role_long <- role_kd_summary %>%
  select(role, avg_kills, avg_deaths) %>%
  pivot_longer(cols = c(avg_kills, avg_deaths), 
               names_to = "stat", values_to = "value")

# Kill and Death Averages by Roles
ggplot(role_long, aes(x = role, y = value, fill = stat)) +
  geom_col(position = "dodge") +
  labs(x = "Role", y = "Average per Game", 
       title = "Kills vs Deaths by Role", fill = "")
```


